---
title: "Estimator Basics"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimators Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/tfestimators
menu:
  main:
    name: "Estimator Basics"
    identifier: "tfestimators-basics"
    parent: "tfestimators-using-tfestimators"
    weight: 20
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

The basic components of the TensorFlow Estimators API include:

- Canned estimators (pre-built implementations of various models).

- Custom estimators (custom model implementations).

- Estimator methods (core methods like `train()`, `predict()`, etc. which work the same for all canned and custom estimators).

- Feature columns (definitions of how features should be transformed during modeling).

- Input functions (sources of data for training and prediction).

In addition, there are APIs that cover more advanced training scenarios:

- Experiments (wrappers around estimators that handle concerns like distributed training, hyperparameter tuning, etc.)

- Run hooks (callbacks for recording context and interacting with training as it proceeds)

Please read our [white paper](http://terrytangyuan.github.io/data/papers/tf-estimators-kdd-paper.pdf) if you are interested in the detailed design of the above components. 

Below we enumerate some of the core functions in each of these componets to provide a high level overview of what's available. See the linked articles for more details on using all of the components together. 


## Canned Estimators

The **tfestimators** package includes a wide variety of canned estimators for common machine learning tasks. Currently, the following canned estimators are available:

| Estimator  | Description |
|---------------------------------------|----------------------------------------------------------------|
| `linear_regressor()`  | Linear regressor model.  |
| `linear_classifier()`  |  Linear classifier model. |
| `dnn_regressor()`  | DNN Regression. |
| `dnn_classifier()`  |  DNN Classification. |
| `dnn_linear_combined_regressor()`  |  DNN Linear Combined Regression. |
| `dnn_linear_combined_classifier()`  | DNN Linear Combined Classification. |

Before you can use an estimator you need to provide an input function and define a set of feature columns. The following two sections cover how to do this.

## Input Functions

Input functions are used to provide data to Estimtor training and prediction methods. The R interface provides several high-level input function implementations for various common R data sources, including:

- Formulas
- Matrices
- Data Frames
- Lists of vectors

For example:

```{r}
input <- input_fn(mtcars, 
                  features = c(drat, cyl), 
                  response = vs,
                  batch_size = 128)
```

The `batch_size` specifies the number of observations to draw for each training step. 

You can also write fully custom input functions that draw data from arbitrary data sources. See the [input functions](input_functions.html) article for additional details.

## Feature Columns

Feature columns speicfy the feature transformations to apply before using data within training or prediction. This code specifies that the `profit` column is a standard numeric variable and that the `profession` column is categorical with 5 buckets:

```{r}
cols <- feature_columns(
  column_numeric("profit"),
  column_categorical_with_identity("profession", num_buckets = 5)
)
```

There are a wide variety of feature column functions available:

| Method  | Description |
|---------------------------------------|----------------------------------------------------------------|
| `column_indicator()`  | Represents multi-hot representation of given categorical column.   |
| `column_numeric()`  |  Represents real valued or numerical features.  |
| `column_embedding()`  | Creates an dense column that converts from sparse, categorical input.   |
| `column_bucketized()`  | Represents discretized dense input.   |
| `column_categorical_weighted()`  | Applies weight values to a categorical column.   |
| `column_categorical_with_vocabulary_list()`  | 	Creates a categorical column with in-memory vocabulary. |
| `column_categorical_with_vocabulary_file()`  | Creates a categorical column with a vocabulary file.   |
| `column_categorical_with_identity()`  | Creates a categorical column that returns identity values.   |
| `column_categorical_with_hash_bucket()`  | Represents sparse feature where ids are set by hashing.   |


See the article on [feature columns](feature_columns.html) for additional details.

## Creating an Estimator

Here's an example of creating a Linear DNN Combined canned Estimator. In creating the estimator we pass the feature columns and other parameters that specifies the layers and architecture of the model. Note that for this particular canned Estimator, it takes two types of feature columns that are used for constructing linear layer and fully connected deep layers separately:

```{r}
classifier <-
	dnn_linear_combined_classifier(
	  linear_feature_columns = linear_feature_columns,
	  dnn_feature_columns = dnn_feature_columns,
	  dnn_hidden_units = c(3L, 3L),
	  dnn_optimizer = "Adagrad"
	)
```

## Training and Prediction

Users can then call `train()` to train the initialized Estimator for a number of steps:

``` r
classifier <- train(classifier, input_fn = custom_input_fn, steps = 2)
```

Once a model is trained, users can use `predict()` that makes predictions on a given input function that represents the inference data source. 

``` r
predictions <- predict(classifier, input_fn = custom_input_fn)
```

Users can also pass a key to `predict_keys` argument in `predict()` that generates different types of predictions, such as probabilities using `prediction_keys()$PROBABILITIES`:

``` r
predictions <- predict(
  classifier,
  input_fn = custom_input_fn,
  predict_keys = prediction_keys()$PROBABILITIES)
```

or logistic using `prediction_keys()$LOGISTIC`:

``` r
predictions <- predict(
  classifier,
  input_fn = custom_input_fn,
  predict_keys = prediction_keys()$LOGISTIC)
```

You can find all the available keys by printing `prediction_keys()`. However, not all keys can be used by different types of estimators. For example, regressors cannot use `PROBABILITIES` as one of the keys.

There's also a function called `coef()` that can be used extract the trained coefficients of a model.

``` r
coefs <- coef(classifier)
```

There are a number of estimator methods which can be used generically with any canned or custom estimator:

To train and generate predictions with estimators you use the following set of generic functions:

| Method  | Description |
|---------------------------------------|----------------------------------------------------------------|
| `train()`  |  Trains a model given training data input_fn. |
| `predict()`  |  Returns predictions for given features. |
| `evaluate()`  |  Evaluates the model given evaluation data input_fn. |
| `export_savedmodel()` | Exports inference graph as a SavedModel into a given directory. |
| `coef()`  |  Get the list coefficients or variables from this model's checkpoint. |





