---
title: "Run Hooks"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run Hooks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/tfestimators
menu:
  main:
    name: "Run Hooks"
    identifier: "tfestimators-run-hooks"
    parent: "tfestimators-advanced"
    weight: 20
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

`SessionRunHooks` are useful to track training, report progress, request early stopping and more. Users can attach an arbitrary number of hooks to an estimator. `SessionRunHooks` use the observer pattern and notify at the following points:

 - when a session starts being used
 - before a call to the `session.run()`
 - after a call to the `session.run()`
 - when the session closed

A `SessionRunHook` encapsulates a piece of reusable/composable computation that can piggyback a call to `MonitoredSession.run()`. A hook can add any ops-or-tensor/feeds to the run call, and when the run call finishes with success gets the outputs it requested. Hooks are allowed to add ops to the graph in `hook.begin()`. The graph is finalized after the `begin()` method is called.

## Built-in Run Hooks

There are a few pre-defined `SessionRunHooks` available, for example:
Run hooks are useful for tracking training, reporting progress, requesting early stopping, and more. Users can attach an arbitrary number of hooks to an estimator. Some built-in run hooks include:

| Method  | Description |
|---------------------------------------|----------------------------------------------------------------|
| `hook_checkpoint_saver()`  |  Saves checkpoints every N steps or seconds. |
| `hook_global_step_waiter()`  | Delay execution until global step reaches to wait_until_step. |
| `hook_history_saver()`  | Saves Metrics History. |
| `hook_logging_tensor()`  | Prints the given tensors once every N local steps or once every N seconds. |
| `hook_nan_tensor()`  | NaN Loss monitor. |
| `hook_progress_bar()`  | Creates and updates progress bar. |
| `hook_step_counter()`  | Steps per second monitor. |
| `hook_stop_at_step()`  | Monitor to request stop at a specified step. |
| `hook_summary_saver()`  | Saves summaries every N steps. |

For example, we can use `hook_progress_bar()` to attach a hook to create and update a progress bar during the model training process. 

```{r eval=FALSE}
fcs <- feature_columns(column_numeric("drat"))
input <- input_fn(mtcars, response = "mpg", features = c("drat", "cyl"), batch_size = 8L)
lr <- linear_regressor(
  feature_columns = fcs
) %>% train(
  input_fn = input,
  steps = 2,
  hooks = list(
    hook_progress_bar()
  ))
```

```
Training 2/2 [======================================] - ETA:  0s - loss: 3136.10
```

Another example is to use `hook_history_saver()` to save the training history every 2 training steps like the following:

```{r eval=FALSE}
lr <- linear_regressor(
  feature_columns = fcs
) %>% train(
  input_fn = input,
  steps = 10,
  hooks = list(
    hook_history_saver(every_n_step = 2)
  ))
```

and then we can use `visualize_metrics("train")` to visualize the metrics saved through training.

## Custom Run Hooks

Users can also create custom run hooks by defining the behaviors of the hook in different phases of a session.

We can implement a custom run hook by defining a list of call back functions as part of `session_run_hook()` initialization. It has the following optional parameters that can be overriden by a custom defined function:

- `begin`: An \R function with signature `function()`, to be called once before using the session.
- `after_create_session`: An \R function with signature `function(session, coord)`, to be called once the new TensorFlow session has been created.
- `before_run`: An \R function with signature `function(run_context)`to be called before a run.
- `after_run`: An \R function with signature  `function(run_context, run_values)` to be called after a run.
- `end`: An \R function with signature `function(session)` to be called at the end of the session.

For example, let's try to implement the `hook_history_saver()` that we showed in previous section. We first initialize a `iter_count` variable to save the current count of the steps being run. We increment the count as part of `after_run()` after each `session.run` calls. Inside `before_run()`, we use the context to access the current losses and save it to a tensor named "losses" so that later we can access it inside `after_run()` via `values$results$losses` that contains the evaluated value of tensor "losses". Finally, we calculate the mean of the raw losses and append it to a global state varibale named "mean_losses_history" with the list of mean losses. 

```{r eval=FALSE}
mean_losses_history <<- NULL
hook_history_saver_custom <- function(every_n_step) {

    iter_count <<- 0

    session_run_hook(

      before_run = function(context) {
        session_run_args(
          losses = context$session$graph$get_collection("losses")
        )
      },
      
      after_run = function(context, values) {
        iter_count <<- iter_count + 1
        print(paste0("Running step: ", iter_count))
        if (iter_count %% every_n_step == 0) {
          raw_losses <- values$results$losses[[1]]
          mean_losses_history <<- c(mean_losses_history, mean(raw_losses))
        }
      }
    )
}
```

Next, we can then attach this hook to our estimator:
```{r eval=FALSE}
lr <- linear_regressor(
  feature_columns = fcs
) %>% train(
  input_fn = input,
  steps = 4,
  hooks = list(
    hook_history_saver_custom(every_n_step = 1)
  ))
```

```
[1] "Running step: 1"
[1] "Running step: 2"
[1] "Running step: 3"
[1] "Running step: 4"
```

We saved the losses history at every step. Let's check the list of losses:

```
> mean_losses_history
[1] 415.8088 452.2128 376.7346 331.6045
```


